
\chapter{Sorting Algorithms}
\Label{cha:sorting}



Many issues in computer science can be exemplified in the field of
sorting algorithms; see e.g.\ \cite{Knuth.1973} for a famous textbook.
Therefore we arrange some of the most common classic
sorting algorithms.
In this chapter, we present algorithms of the \cxx Standard
Library \cite[\S 28.7.1]{cxx-17-draft} that are related to the task
of sorting a linear array.

Following \cite{Sellibitze.2014}, we have also used (\isoc rephrasings of) 
functions from the \cxx Standard Library as far as possible to implement the different
algorithmic approaches.

\begin{itemize}
\item \issorted in \S\ref{sec:issorted} is an algorithm that checks if a given
array is already in increasing order.

\item \partialsort in \S\ref{sec:partialsort} rearranges a given
array into two parts. All elements in the first part are less or equal than
those of the second part. Moreover, while the first part is sorted,
the order of elements in the second part is unspecified.

\item \bubblesort in \S\ref{sec:bubblesort} describes a simple, well-known and 
      sorting algorithm.\footnote{
   See \url{https://en.wikipedia.org/wiki/Bubble_sort}
}

\item
\selectionsort in \S\ref{sec:selectionsort} presents the classic
\emph{selection sort} algorithm.\footnote{
   See \url{https://en.wikipedia.org/wiki/Selection_sort}
}

\item \insertionsort in \S\ref{sec:insertionsort} the also well-known
\emph{insertion sort} algorithm.\footnote{
   See \url{https://en.wikipedia.org/wiki/Insertion_sort}
}

\item \heapsort in \S\ref{sec:heapsort} describes the quite efficient \emph{heap sort},
which relies on the algorithms presented in Chapter~\ref{cha:heap}.\footnote{
   See \url{https://en.wikipedia.org/wiki/Heapsort}
}

\item \merge in \S\ref{sec:merge} the \emph{merge} algorithm from \emph{merge sort}.\footnote{
  See \url{https://en.wikipedia.org/wiki/Merge_sort}
}
\end{itemize}

While \heapsort achieves a run-time complexity upper bound of 
${\mathcal O}(n \cdot \log(n))$ due to the efficiency of the heap
data structure,
both \selectionsort and \insertionsort
need ${\mathcal O}(n^2)$ in the average case,
and also in the worst case.

Note that the \inl{sort} algorithm from the \cxx Standard Library
is not handled here because it typically relies on \emph{introspection sort}
which is sophisticated mix of various classic algorithms.\footnote{
  See \url{https://en.wikipedia.org/wiki/Introsort}
}
In future releases we plan to handle the more algorithms related sorting.

\clearpage

The sorting algorithms in this chapter essentially share the following contract;
it is their implementations that differ fundamentally.

\begin{lstlisting}[style = acsl-block]
        /*@
           requires valid: \valid(a + (0..n-1));

           assigns  a[0..n-1];

           ensures increasing:  Increasing(a, n);
           ensures reorder:     MultisetReorder{Old, Here}(a, n);
        */
        void xxx_sort(value_type* a, size_type n);
\end{lstlisting}

As mentioned in the introduction of Chapter~\ref{cha:binary-search},
we use the predicate \logicref{Increasing} in the contracts of our
sorting algorithms but often resort to the simpler predicate
\logicref{WeaklyIncreasing} in the loop invariants and assertions.
In order to conclude that the desired postcondition \inl{Increasing(a, n)} holds,
we rely on lemma \logicref{WeaklyIncreasingIncreasing}.


%\clearpage

\input{sorting/is_sorted}
\input{sorting/partial_sort}
\input{sorting/bubble_sort}
\input{sorting/selection_sort}
\input{sorting/insertion_sort}
\input{sorting/heap_sort}
\input{sorting/merge}
